Deep mutational scanning of hemagglutinin helps predict evolutionary fates of human H3N2 influenza variants

### 2018-06-24 paper/manuscript.tex
Remove un-scaled Perth logoplot supp figure

We next sought to quantify the correlation between a mutation's experimentally measured effect and the maximum frequency it attained during natural evolution.
To calculate a given mutation's effect, we simply took the logarithm of the ratio of the preferences for the mutant and wildtype amino acids at that site.
To minimize effects related to the genetic background of the strain used in the experiment, we excluded mutations closely related to the experimental strain itself and partitioned the remaining mutations into 1,022 mutations pre-dating and 299 mutations post-dating the Perth/2009 strain (Figure~\ref{suppfig:tree}).
We additionally excluded mutations from the post-Perth partition that were sampled in 2014 or after that have not had enough time for their evolutionary fates to be fully resolved.
We used these pre-Perth and post-Perth partitions to test the utility of our measurements for both post-hoc and prospective analyses, respectively.
We quantified the relationship between mutational effects and maximum mutation frequencies in the H3N2 phylogeny via Spearman rank correlation (Figure~\ref{fig:muteffect_maxfreq}A).
In both pre-Perth and post-Perth time periods, we found a modest, but statistically significant relationship between mutational effect and maximum mutation frequency (pre-Perth $\rho = 0.17$, post-Perth $\rho = 0.14$). 
The similar effect sizes for both the pre- and post-Perth partitions shows that our experimental measurements can help explain the evolutionary fates of mutations in strains that post-date the experimentally studied strain, as well as to retrospectively analyze mutations that precede the experimental strain.

@@

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.8\textwidth]{figs/supp_WSNprefs_logoplot/supp_logoplots.pdf}}
\caption{\label{suppfig:WSNprefs_logoplot}
{\bf The site-specific amino-acid preferences of the WSN/1933 H1 HA.}
\comment{Get rid of unscaled Perth/2009 in this figure. It doesn't make sense to show them here. I've removed from legend; remove from the actual figure}
The amino-acid preferences of the WSN/1933 H1 HA from~\cite{doud2016accurate} after taking the average of the experimental replicates and re-scaling~\cite{hilton2017phydms} by a stringency parameter of 2.05 (see \url{https://github.com/jbloomlab/dms_tools2/blob/master/examples/Doud2016/analysis_notebook.ipynb}).
The overlays show the same information as in Figure~\ref{fig:logoplot} (domain and wildtype amino acid).
Note that \cite{doud2016accurate} used libraries in which all codons were mutagenized \emph{except} for the one encoding N-terminal methionine.
Therefore, no data is shown for the first codon in the gene.
The sites are in H3 numbering.
}
\end{suppfigure}

>>>>

We next sought to quantify the correlation between a mutation's experimentally measured effect and the maximum frequency it attained during natural evolution.
To calculate a given mutation's effect, we simply took the logarithm of the ratio of the preferences for the mutant and wildtype amino acids at that site.
To minimize effects related to the genetic background of the strain used in the experiment, we excluded mutations closely related to the experimental strain itself and partitioned the remaining mutations into 1,022 mutations pre-dating and 299 mutations post-dating the Perth/2009 strain (Figure~\ref{suppfig:tree}).
We additionally excluded mutations from the post-Perth partition that were sampled in 2014 or after that have not had enough time for their evolutionary fates to be fully resolved.
We used these pre-Perth and post-Perth partitions to test the utility of our measurements for both post-hoc and prospective analyses, respectively.
We quantified the relationship between mutational effects and maximum mutation frequencies in the H3N2 phylogeny via Spearman rank correlation (Figure~\ref{fig:muteffect_maxfreq}A).
In both pre-Perth and post-Perth time periods, we found a modest, but statistically significant relationship between mutational effect and maximum mutation frequency (pre-Perth $\rho = 0.17$, post-Perth $\rho = 0.15$). 
The similar effect sizes for both the pre- and post-Perth partitions shows that our experimental measurements can help explain the evolutionary fates of mutations in strains that post-date the experimentally studied strain, as well as to retrospectively analyze mutations that precede the experimental strain.

@@

\begin{suppfigure}[H]
\centerline{\includegraphics[width=0.9\textwidth]{figs/supp_WSNprefs_logoplot/WSN-rescaled_prefs.pdf}}
\caption{\label{suppfig:WSNprefs_logoplot}
{\bf The site-specific amino-acid preferences of the WSN/1933 H1 HA.}
The amino-acid preferences of the WSN/1933 H1 HA from~\cite{doud2016accurate} after taking the average of the experimental replicates and re-scaling~\cite{hilton2017phydms} by a stringency parameter of 2.05 (see \url{https://github.com/jbloomlab/dms_tools2/blob/master/examples/Doud2016/analysis_notebook.ipynb}).
The overlays show the same information as in Figure~\ref{fig:logoplot} (domain and wildtype amino acid).
Note that \cite{doud2016accurate} used libraries in which all codons were mutagenized \emph{except} for the one encoding N-terminal methionine.
Therefore, no data is shown for the first codon in the gene.
The sites are in H3 numbering.
}
\end{suppfigure}
